### 2014-10-09 writing/AAAI SSS '15/join.tex
Attempting to resolve conflicts, shrunk figs
More compact join section

\begin{table}[tp]
  \centering
  \setlength{\tabcolsep}{15pt}
  \renewcommand{\arraystretch}{1.1}
  \begin{tabular}{l c l l} 
    \toprule
    Name & Symbol & Set-theoretic definition & Example \\ 
    \midrule
    entailment         & $x \natfor y$   & $x \subset y$ & \ii{turtle, reptile}  \\ 
    reverse entailment & $x \natrev y$   & $x \supset y$ & \ii{reptile, turtle}  \\ 
    equivalence        & $x \nateq y$    & $x = y$       & \ii{couch, sofa} \\ 
    alternation        & $x \natalt y$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{turtle, warthog} \\ 
    negation           & $x \natneg y$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$    & \ii{able, unable} \\
    cover              & $x \natcov y$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal, non-turtle} \\ 
    independence       & $x \natind y$   & (else) & \ii{turtle, pet}\\
    \bottomrule
  \end{tabular}
  \caption{The seven natural logic relations of \cite{maccartney2009extended}. 
    $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, 
    and the relation $\natind$ applies whenever none of the other six do.} %, including when there 
    %is insufficient knowledge to choose a label.}
  \label{b-table}
\end{table}

If a model is to learn the behavior of a relational logic like the one
presented here from a finite amount of data, it must learn to deduce new
relations from seen relations in a sound manner. The simplest such
deductions involve atomic statements using the relations in
Table~\ref{b-table}. For instance, given that $a \natrev b$ and $b
\natrev c$, one can conclude that $a \natrev c$, by basic
set-theoretic reasoning (transitivity of $\natrev$). Similarly, from
$a \natfor b$ and $b \natneg c$, it follows that $a \natalt c$.  The
full set of sound inferences of this form is summarized in
Table~\ref{tab:jointable}; cells containing a dot correspond to pairs
of relations for which no valid inference can be drawn in our logic.

@@

\begin{table}[htp]
  \centering  
  \setlength{\arraycolsep}{8pt}
  \renewcommand{\arraystretch}{1.1}
  \newcommand{\UNK}{\cdot}  
  $\begin{array}[t]{c@{ \ }|*{7}{c}|}
    %\hline
    \multicolumn{1}{c}{}
             & \nateq     & \natfor     & \natrev     & \natneg    & \natalt     & \natcov     & \multicolumn{1}{c}{\natind} \\
    \cline{2-8}
    \nateq  & \nateq &   \natfor &  \natrev &  \natneg &   \natalt &  \natcov &  \natind \\
    \natfor & \natfor &  \natfor &  \UNK &  \natalt &   \natalt &  \UNK &  \UNK \\
    \natrev & \natrev &  \UNK &  \natrev &  \natcov &   \UNK &  \natcov &  \UNK \\
    \natneg & \natneg &  \natcov &  \natalt &  \nateq &    \natrev &  \natfor &  \natind \\
    \natalt & \natalt &  \UNK &  \natalt &  \natfor &   \UNK &  \natfor &  \UNK \\
    \natcov & \natcov &  \natcov &  \UNK &  \natrev &   \natrev &  \UNK &  \UNK \\
    \natind & \natind & \UNK &  \UNK &  \natind &  \UNK &  \UNK &  \UNK \\
    \cline{2-8}
  \end{array}$
  \caption{Inference path from premises $a\,R\,b$ (row) and $b\,S\,c$ (column) to the relation that holds between $a$ and $c$, if any.  These inferences are based on basic set-theoretic truths about the meanings of the underlying relations as described in Table~\ref{b-table}. We assess our models' ability to reproduce such inferential paths.}
  \label{tab:jointable}
\end{table}

\paragraph{Experiments}
To test our models' ability to learn these inferential patterns, we
create small boolean structures for our logic in which terms denote
sets of entities from a small domain.  Figure~\ref{lattice-figure}
depicts a structure of this form. The lattice gives the full model,
for which all the statements on the right are valid. We divide these
statements evenly into training and test sets, and remove from the
test set those statements which cannot be proven from the training
statements.
% using the logic depicted in Figure~\ref{lattice-figure}.

% We test the model's ability to learn this behavior by creating
% artificial data sets of terms which represent sets of numbers. Since
% MacCartney and Manning's set of relations hold between sets as well as
% between sentences, we can use the underlying set structure to generate
% the all of the relations that hold between any pair of these terms, as
% in Figure \ref{lattice-figure}. We train the model defined above on a
% subset of these relations, but rather then presenting the model with a
% pair of tree-structured sentences as inputs, simply present it with
% two single terms, each of which corresponds to a single vector in the
% (randomly initialized) vocabulary matrix $V$, ensuring that the model
% has no information about the terms being compared except the relations
% between them.

In our experiments, we create 80 randomly generated sets drawing from
a domain of seven elements. This yields a data set consisting of
6400 statements about pairs of formulae. 3200 of these pairs are
chosen as a test set, and that test set is further reduced to the 2960
examples that can be provably derived from the training data.

We trained versions of both the RNN model and the RNTN model on these
data sets. In both cases, the models were implemented exactly as
described in Section~\ref{methods}, but since the items being compared
are single terms rather than full tree structures, the composition
layer was not involved, and the two models differed only in which
layer function was used for the comparison layer. We simply present
the models with two single terms, each of which corresponds to a
single vector in the (randomly initialized) vocabulary matrix $V$,
thereby ensuring that the model has no information about the terms
being compared except the relations between them.

\paragraph{Results} 
We found that the RNTN model worked best with 11-dimensional vector
representations for the 80 sets and a 90-dimensional feature vector
for the classifier, though the performance of the model was not highly
sensitive to either dimensionality parameter in any of the experiments
discussed here. This model was able to correctly label 99.4\% (94.3
macroaveraged F1) of the provably derivable test examples, and 95.4\%
(92.2 macro-F1) of the remaining test examples. The simpler RNN model worked
best with 11 and 75 dimensions, respectively, but was able to achieve
accuracies of only 89.0\% (80.7 macro-F1) and 62.5\% (40.7 macro-F1),
respectively. Training accuracy was 99.8\% for the RNTN and 95.4\% for
the RNN.

\begin{figure}[tp]
  \centering
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \newcommand{\labelednode}[4]{\put(#1,#2){\oval(1.5,1)}\put(#1,#2){\makebox(0,0){$\begin{array}{c}#3\\\{#4\}\end{array}$}}}
    \setlength{\unitlength}{1cm}
    \begin{picture}(5,5.5)
      \labelednode{2.50}{5}{}{1,2,3}

@@

      \labelednode{2.5}{0.5}{}{}
    \end{picture}
    \caption{Simple boolean structure. The letters name the sets. Not all sets have names, and
    some sets have multiple names, so that learning $\nateq$ is non-trivial.}
  \end{subfigure}
  \qquad
  \begin{subfigure}[t]{0.43\textwidth}
    \centering
    \setlength{\tabcolsep}{12pt}
    \begin{tabular}[b]{c  c}
      \toprule
      Train & Test \\
      \midrule
                    & $b \nateq b$ \\
      $b \natcov c$ &               \\
                    & $b \natcov d$ \\
                    & \strikeout{$b \natrev e$} \\
      $c \natcov d$ &               \\
      $c \natrev e$ &               \\
                    & \strikeout{$c \nateq f$} \\
      $c \natrev g$ &               \\ 
                    & $e \natfor b$ \\
      $e \natfor c$ &               \\[-1ex]
      $\vdots$      & $\vdots$ \\
      \bottomrule
    \end{tabular}

    \caption{A train/test split of the atomic statements about the
      model.  Test statements not provable from the training data are
      crossed out.}
  \end{subfigure}  
  \caption{Small example structure and data for learning relation composition.}
  \label{lattice-figure}
\end{figure} 
% Note: None of these test examples is derivable from the shown training data, 
% but we suggest that there is additional training data, so we can cross lines
% out willy-nilly without fear.

@@

These results are fairly straightforward to interpret. The RNTN model
was able to accurately encode the relations between the terms in the
geometric relations between their vectors, and was able to then use
that information to recover relations that were not overtly included
in the training data. In contrast, the RNN model was able to achieve
this behavior only incompletely. It is possible but not likely that it
could be made to find a good solution with further optimization on
different learning algorithms, or that it would do better on a larger
universe of sets, for which there would be a larger set of training
data to learn from, but the RNTN is readily able to achieve these
effects in the setting discussed here.

>>>>

\begin{table}[tp]
  \centering\small
  \setlength{\tabcolsep}{15pt}
  \renewcommand{\arraystretch}{1.1}
  \begin{tabular}{l c l l} 
    \toprule
    Name & Symbol & Set-theoretic definition & Example \\ 
    \midrule
    entailment         & $x \natfor y$   & $x \subset y$ & \ii{turtle, reptile}  \\ 
    reverse entailment & $x \natrev y$   & $x \supset y$ & \ii{reptile, turtle}  \\ 
    equivalence        & $x \nateq y$    & $x = y$       & \ii{couch, sofa} \\ 
    alternation        & $x \natalt y$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{turtle, warthog} \\ 
    negation           & $x \natneg y$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$    & \ii{able, unable} \\
    cover              & $x \natcov y$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal, non-turtle} \\ 
    independence       & $x \natind y$   & (else) & \ii{turtle, pet}\\
    \bottomrule
  \end{tabular}
  \caption{The seven natural logic relations of \cite{maccartney2009extended}. 
    $\mathcal{D}$ is the universe of possible objects of the same type as those being compared, 
    and the relation $\natind$ applies whenever none of the other six do.} %, including when there 
    %is insufficient knowledge to choose a label.}
  \label{b-table}
\end{table}

If a model is to learn the behavior of a relational logic like the one
presented here from a finite amount of data, it must learn to deduce new
relations from seen relations in a sound manner. The simplest such
deductions involve atomic statements using the relations in
Table~\ref{b-table}. For instance, given that $a \natrev b$ and $b
\natrev c$, one can conclude that $a \natrev c$, by basic
set-theoretic reasoning (transitivity of $\natrev$). Similarly, from
$a \natfor b$ and $b \natneg c$, it follows that $a \natalt c$.  The
full set of sound inferences of this form is summarized in
Figure~\ref{tab:jointable}; cells containing a dot correspond to pairs
of relations for which no valid inference can be drawn in our logic.

@@

\paragraph{Experiments}
To test our models' ability to learn these inferential patterns, we
create boolean structures in which terms denote sets of entities from
a small domain (e.g., Figure~\ref{lattice-figure}), employ logical
deduction to identify the valid statements, divide those into train
and test sets, and remove from the test set those statements which
cannot be proven from the training statements
(Figure~\ref{unprovable}).
% using the logic depicted in Figure~\ref{lattice-figure}.
%
%
% We test the model's ability to learn this behavior by creating
% artificial data sets of terms which represent sets of numbers. Since
% MacCartney and Manning's set of relations hold between sets as well as
% between sentences, we can use the underlying set structure to generate
% the all of the relations that hold between any pair of these terms, as
% in Figure \ref{lattice-figure}. We train the model defined above on a
% subset of these relations, but rather then presenting the model with a
% pair of tree-structured sentences as inputs, simply present it with
% two single terms, each of which corresponds to a single vector in the
% (randomly initialized) vocabulary matrix $V$, ensuring that the model
% has no information about the terms being compared except the relations
% between them.
%
In our experiments, we create 80 randomly generated sets drawn from a
domain of seven elements. This yields 6400 statements about pairs of
formulae, of which 3200 are chosen as a test set, and that test set is
further reduced to the 2960 examples that can be provably derived from
the training data. We trained both the NN model and the NTN model on
these data
sets% . In both cases, the models were implemented exactly as
% described in Section~\ref{methods}, but since the items being compared
% are single terms rather than full tree structures, the composition
% layer was not involved, and the two models differed only in which
% layer function was used for the comparison layer. We simply present
% the models with two single terms, each of which corresponds to a
% single vector in the (randomly initialized) vocabulary matrix $V$,
% thereby ensuring that the model has no information about the terms
% being compared except the relations between them.

\paragraph{Results} 
We found that the NTN model worked best with 11-dimensional vector
representations for the 80 sets and a 90-dimensional feature vector
for the classifier, though the performance of the model was not highly
sensitive to either dimensionality parameter in any of the experiments
discussed here. This model was able to correctly label 99.4\% (94.3
macroaveraged F1) of the provably derivable test examples, and 95.4\%
(92.2 macro-F1) of the remaining test examples. The simpler NN worked
best with 11 and 75 dimensions, respectively, but was able to achieve
accuracies of only 89.0\% (80.7 macro-F1) and 62.5\% (40.7 macro-F1),
respectively. Training accuracy was 99.8\% for the NTN and 95.4\% for
the NN.

\begin{figure}[tp]
  \centering
  \begin{subfigure}[b]{0.42\textwidth}
    \centering  
    \setlength{\arraycolsep}{8pt}
    \renewcommand{\arraystretch}{1.1}
    \newcommand{\UNK}{\cdot}  
    \resizebox{2.2in}{!}{$\begin{array}[c]{c@{ \ }|*{7}{c}|}
        % \hline
        \multicolumn{1}{c}{}
        & \nateq     & \natfor     & \natrev     & \natneg    & \natalt     & \natcov     & \multicolumn{1}{c}{\natind} \\
        \cline{2-8}
        \nateq  & \nateq &   \natfor &  \natrev &  \natneg &   \natalt &  \natcov &  \natind \\
        \natfor & \natfor &  \natfor &  \UNK &  \natalt &   \natalt &  \UNK &  \UNK \\
        \natrev & \natrev &  \UNK &  \natrev &  \natcov &   \UNK &  \natcov &  \UNK \\
        \natneg & \natneg &  \natcov &  \natalt &  \nateq &    \natrev &  \natfor &  \natind \\
        \natalt & \natalt &  \UNK &  \natalt &  \natfor &   \UNK &  \natfor &  \UNK \\
        \natcov & \natcov &  \natcov &  \UNK &  \natrev &   \natrev &  \UNK &  \UNK \\
        \natind & \natind & \UNK &  \UNK &  \natind &  \UNK &  \UNK &  \UNK \\
        \cline{2-8}
      \end{array}$}
    \caption{Inference path from premises $a\,R\,b$ (row) and $b\,S\,c$ (column) to the relation that holds between $a$ and $c$, if any.  These inferences are based on basic set-theoretic truths about the meanings of the underlying relations as described in Table~\ref{b-table}. We assess our models' ability to reproduce such inferential paths.}
    \label{tab:jointable}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
    \newcommand{\labelednode}[4]{\put(#1,#2){\oval(1.5,1)}\put(#1,#2){\makebox(0,0){$\begin{array}{c}#3\\\{#4\}\end{array}$}}}
    \setlength{\unitlength}{1cm}
    \resizebox{1.5in}{!}{\begin{picture}(5,5.5)
      \labelednode{2.50}{5}{}{1,2,3}

@@

      \labelednode{2.5}{0.5}{}{}
    \end{picture}}
    \caption{Simple boolean structure. The letters name the sets. Not all sets have names, and
    some sets have multiple names, so that learning $\nateq$ is non-trivial.}\label{lattice-figure}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.2\textwidth}
    \centering
    \setlength{\tabcolsep}{12pt}
    \resizebox{1.05in}{!}{\begin{tabular}[c]{c  c}
      \toprule
      Train & Test \\
      \midrule
                    & $b \nateq b$ \\
      $b \natcov c$ &               \\
                    & $b \natcov d$ \\
                    & \strikeout{$b \natrev e$} \\
      $c \natcov d$ &               \\
      $c \natrev e$ &               \\
                    & \strikeout{$c \nateq f$} \\
      $c \natrev g$ &               \\ 
                    & $e \natfor b$ \\
      $e \natfor c$ &               \\[-1ex]
      $\vdots$      & $\vdots$ \\
      \bottomrule
    \end{tabular}}

    \caption{A train/test split of the atomic statements about the
      model.  Test statements not provable from the training data are
      crossed out.}\label{unprovable}
  \end{subfigure}  
  \caption{Small example structure and data for learning relation composition.}
  \label{exp1}
\end{figure} 
% Note: None of these test examples is derivable from the shown training data, 
% but we suggest that there is additional training data, so we can cross lines
% out willy-nilly without fear.

@@

These results are fairly straightforward to interpret. The NTN model
was able to accurately encode the relations between the terms in the
geometric relations between their vectors, and was able to then use
that information to recover relations that were not overtly included
in the training data. In contrast, the NN was able to achieve this
behavior only incompletely. It is possible but not likely that it
could be made to find a good solution with further optimization on
different learning algorithms, or that it would do better on a larger
universe of sets, for which there would be a larger set of training
data to learn from, but the NTN is readily able to achieve these
effects in the setting discussed here.

### 2014-10-09 writing/AAAI SSS '15/methods.tex
More compact join section

We build a model to learn embeddings using the method of
\citet{Bowman:Potts:Manning:2014}, which is centered on the problem of
labeling a pair of words or sentences with one of a small set of
logical relations. The architecture of the model that we use, which is
limited to only pairs of single symbols (such as words), is depicted
in Figure~\ref{sample-figure}. The model represents the two input
symbols as embeddings, which are fed into a comparison function based
on one of two types of neural network layer functions to produce a
representation for the relationship between the two symbols. This
representation is then fed into a simple softmax classifier which
outputs a distribution over possible labels. The entire network,
including the embeddings, are trained through backpropagation with
AdaGrad \cite{duchi2011adaptive}.

@@

For a comparison funtion, we evaluate versions of the model with both a plain neural
network (NN) layer function and a neural tensor network (NTN) layer function
proposed in \citet{chen2013learning}, which adds a third order tensor parameter
to mediate multiplicative interactions between the two inputs. A leaky ReLU
nonlinearity \cite{maasrectifier} is applied to the output of either
layer function.
%
\begin{gather} \label{rnn}
\vec{y}_{\textit{NN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}) ~~~~~~~~~~~~~~~ \vec{y}_{\textit{NTN}} = f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)} + \mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b})
\end{gather} 
%

This model differs from that of \citet{Bowman:Potts:Manning:2014} in two ways. Because 
the inputs are single symbols, there is no need for the composition functions
which are used in that (and prior) work. Also, for our experiment on 
WordNet data, we introduce an additional neural network layer between the embedding input
and the comparison function, which is meant to facilitate initializing the embeddings
themselves from pretrained vectors, and was found to help peformance in that setting.

>>>>

We build a model to learn embeddings using the method of
\citet{Bowman:Potts:Manning:2014}, which is centered on the problem of
labeling a pair of words or sentences with one of a small set of
logical relations. The architecture of the model that we use, which is
limited to only pairs of single symbols (such as words), is depicted
in Figure~\ref{sample-figure}. The model represents the two input
symbols as embeddings, which are fed into a comparison function based
on one of two types of neural network layer functions to produce a
representation for the relationship between the two symbols. This
representation is then fed into a softmax classifier, which outputs a
distribution over possible labels. The entire network, including the
embeddings, are trained through backpropagation with AdaGrad
\cite{duchi2011adaptive}.

@@

For a comparison funtion, we evaluate versions of the model with both
a plain NN and the NTN layer function proposed in
\citet{chen2013learning}, which adds a third order tensor parameter to
mediate multiplicative interactions between the two inputs. A leaky
ReLU nonlinearity \cite{maasrectifier} is applied to the output of
either layer function.
%
\begin{gather} \label{rnn}
\vec{y}_{\textit{NN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}) ~~~~~~~~~~~~~~~ \vec{y}_{\textit{NTN}} = f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)} + \mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b})
\end{gather} 
%

This model differs from that of \citet{Bowman:Potts:Manning:2014} in
two ways. Because the inputs are single symbols, there is no need for
the composition functions used in that (and prior) work. Also, for our
experiment on WordNet data, we introduce an additional neural network
layer between the embedding input and the comparison function, which
facilitates initializing the embeddings themselves from pretrained
vectors and was found to help peformance in that setting.
